---
title: "voting_data_attempt"
author: "Niels Alderliesten"
date: "2026-09-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(RSiena)
library(igraph)
library(tidyverse)
library(igraph)
library(dutchparl)
library(R.filesets)
```

## Some initial exploration with other data origins

### The Dutch Parliamentary Behaviour Dataset:

<https://research.rug.nl/en/publications/the-dutch-parliamentary-behaviour-dataset/>

```{r }
votes <- loadRDS("C:/Users/niels/Downloads/RMSCS/Year 2/Social Networks/labjournal_social_networks/data/votes.rds")

partyInfo<- votes[[7]]

party_name_short <- partyInfo[[4]]

party_name_short_df <- as.data.frame(party_name_short)

View(party_name_short_df)

votesame(examplevotes)

?votesame
```

### Setting up a matrix from scratch, subsequently removing NAs

```{r}
testm <- matrix(NA, nrow=4, ncol=4)
testm[1,1] <- 0
testm[1,2] <- 1
testm[1,3] <- 0
testm[1,4] <- 1
testm[2,1] <- 1
testm[2,4] <- 0
testm[3,4] <- 1
testm[3,3] <- 1
testm[3,2] <- 0

missingrow <- rowSums(is.na(testm)) == nrow(testm)
missingcol <- colSums(is.na(testm)) == ncol(testm)

totalmissing <- which(missingrow & missingcol) #Welke ontbreken compleet, welke zou je dus weg kunnen halen.

which(missingrow)
testm
testm[which(missingrow),] <- 0
testm[, which(missingcol)] <- 0

missingrow
missingcol
#Voor RSienna hij was 1.0 = speciale missing waarde. Hij kan wel genomineerd worden maar hij was er niet.
#NA blijft als hij er volgende keer wel bij is

```

### Attempting to see if I can webscrape ONE! value

```{r}
#schoof1_url <- "https://www.partijgedrag.nl/party-likeness"
#schoof1 <- rvest::read_html(schoof1_url)

#schoof1_thingy <- schoof1 %>% rvest::html_elements("body") %>% html_elements("div") %>% html_children() %>%
#  html_elements("main")


#schoof1_thingy[2] %>% html_children %>% html_elements("main") %>% html_children() %>% html_children()
  
  
  

#  html_elements("table") %>%
#  html_elements("tbody") %>% html_children

#html_elements(schoof1, html\body\div[2]\div\main\div\div[3]\div\div[2]\div\table\tbody)

```

Tot nu toe nog niet gelukt helaas.

## Own Data

Data personally gathered from openkamer.org

<https://www.openkamer.org/stats/stemmingen/partijen/>

As well as from the Chapell Hill Data

### List object

First, we create a list that will later function as a collection where we can place all data Note: It is not necessary to add all the sub-items already. These can also be added during the process. However I liked it, in order to keep overview of what needed to be done.

```{r}
partijenstemgedrag <- list(balkenende4 = list(NA, NA, NA, NA, NA, NA),
                           rutte1 = list(NA, NA, NA, NA, NA, NA), 
                           rutte2 = list(NA, NA, NA, NA, NA, NA), 
                           rutte3 = list(NA, NA, NA, NA, NA, NA),
                           rutte4 = list(NA, NA, NA, NA, NA, NA), 
                           schoof1 = list(NA, NA, NA, NA, NA, NA))

names(partijenstemgedrag$balkenende4) <- c("votesbalk", "originvotesbalk", "adjvotesbalk", "statsbalk", "weights1balk", "weights2balk")

names(partijenstemgedrag$rutte1) <- c("votesrutte1", "originvotesrutte1", "adjvotesrutte1", "statsrutte1", "weights1rutte1", "weights2rutte1")

names(partijenstemgedrag$rutte2) <- c("votesrutte2", "originvotesrutte2", "adjvotesrutte2", "statsrutte2", "weights1rutte2", "weights2rutte2")

names(partijenstemgedrag$rutte3) <- c("votesrutte3", "originvotesrutte3", "adjvotesrutte3", "statsrutte3", "weights1rutte3", "weights2rutte3")

names(partijenstemgedrag$rutte4) <- c("votesrutte4", "originvotesrutte4", "adjvotesrutte4", "statsrutte4", "weights1rutte4", "weights2rutte4")

names(partijenstemgedrag$schoof1) <- c("votesschoof1", "originvotesschoof1", "adjvotesschoof1", "statsschoof1", "weights1schoof1", "weights2schoof1")

```

### Votes

First, we download all the votes dataframes

```{r}
library(readxl)

df_balkenende4 <- as.data.frame(readxl::read_excel("C:/Users/niels/Downloads/RMSCS/Year 2/Social Networks/labjournal_social_networks/data/partijgedrag_data.xlsx", sheet = 1))

df_rutte1 <- as.data.frame(readxl::read_excel("C:/Users/niels/Downloads/RMSCS/Year 2/Social Networks/labjournal_social_networks/data/partijgedrag_data.xlsx", sheet = 3))

df_rutte2 <- as.data.frame(readxl::read_excel("C:/Users/niels/Downloads/RMSCS/Year 2/Social Networks/labjournal_social_networks/data/partijgedrag_data.xlsx", sheet = 5))

df_rutte3 <- as.data.frame(readxl::read_excel("C:/Users/niels/Downloads/RMSCS/Year 2/Social Networks/labjournal_social_networks/data/partijgedrag_data.xlsx", sheet = 7))

df_rutte4 <- as.data.frame(readxl::read_excel("C:/Users/niels/Downloads/RMSCS/Year 2/Social Networks/labjournal_social_networks/data/partijgedrag_data.xlsx", sheet = 9))

df_schoof1 <- as.data.frame(readxl::read_excel("C:/Users/niels/Downloads/RMSCS/Year 2/Social Networks/labjournal_social_networks/data/partijgedrag_data.xlsx", sheet = 11))

```

Subsequently, we clean up the votes dataframes and make them ready (in the format of adjacency matrix)

```{r, warning=F}

rownames(df_balkenende4) <- df_balkenende4[,1]

df_balkenende4 <- df_balkenende4[,-1]

df_balkenende4[] <- lapply(df_balkenende4, as.numeric)

#Let's keep this one apart as well, just in case
df_balkenende4_ruw <- df_balkenende4

df_balkenende4[] <- lapply(df_balkenende4, function(x) ifelse(x>.5, 1, x))
df_balkenende4[] <- lapply(df_balkenende4, function(x) ifelse(x<=.5, 0, x))

```

```{r, warning=F}

rownames(df_rutte1) <- df_rutte1[,1]

df_rutte1 <- df_rutte1[,-1]

df_rutte1[] <- lapply(df_rutte1, as.numeric)

df_rutte1_ruw <- df_rutte1

df_rutte1[] <- lapply(df_rutte1, function(x) ifelse(x>.5, 1, x))
df_rutte1[] <- lapply(df_rutte1, function(x) ifelse(x<=.5, 0, x))

```

```{r, warning=F}

rownames(df_rutte2) <- df_rutte2[,1]

df_rutte2 <- df_rutte2[,-1]

df_rutte2[] <- lapply(df_rutte2, as.numeric)

df_rutte2_ruw <- df_rutte2

df_rutte2[] <- lapply(df_rutte2, function(x) ifelse(x>.5, 1, x))
df_rutte2[] <- lapply(df_rutte2, function(x) ifelse(x<=.5, 0, x))

```

```{r, warning=F}

rownames(df_rutte3) <- df_rutte3[,1]

df_rutte3 <- df_rutte3[,-1]

df_rutte3[] <- lapply(df_rutte3, as.numeric)

df_rutte3_ruw <- df_rutte3

df_rutte3[] <- lapply(df_rutte3, function(x) ifelse(x>.5, 1, x))
df_rutte3[] <- lapply(df_rutte3, function(x) ifelse(x<=.5, 0, x))

```

```{r, warning=F}

rownames(df_rutte4) <- df_rutte4[,1]

df_rutte4 <- df_rutte4[,-1]

df_rutte4[] <- lapply(df_rutte4, as.numeric)

df_rutte4_ruw <- df_rutte4

df_rutte4[] <- lapply(df_rutte4, function(x) ifelse(x>.5, 1, x))
df_rutte4[] <- lapply(df_rutte4, function(x) ifelse(x<=.5, 0, x))

```

```{r, warning=F}

rownames(df_schoof1) <- df_schoof1[,1]

df_schoof1 <- df_schoof1[,-1]

df_schoof1[] <- lapply(df_schoof1, as.numeric)

df_schoof1_ruw <- df_schoof1

df_schoof1[] <- lapply(df_schoof1, function(x) ifelse(x>.5, 1, x))
df_schoof1[] <- lapply(df_schoof1, function(x) ifelse(x<=.5, 0, x))

```

Fill in the slots for votes!

```{r}

#balkenende4
partijenstemgedrag$balkenende4$votesbalk <- df_balkenende4
partijenstemgedrag$balkenende4$originvotesbalk <- df_balkenende4_ruw

#Rutte1

partijenstemgedrag$rutte1$votesrutte1 <- df_rutte1
partijenstemgedrag$rutte1$originvotesrutte1 <- df_rutte1_ruw

#Rutte2
partijenstemgedrag$rutte2$votesrutte2 <- df_rutte2
partijenstemgedrag$rutte2$originvotesrutte2 <- df_rutte2_ruw

#Rutte3
partijenstemgedrag$rutte3$votesrutte3 <- df_rutte3
partijenstemgedrag$rutte3$originvotesrutte3 <- df_rutte3_ruw

#Rutte4
partijenstemgedrag$rutte4$votesrutte4 <- df_rutte4
partijenstemgedrag$rutte4$originvotesrutte4 <- df_rutte4_ruw

#Schoof1
partijenstemgedrag$schoof1$votesschoof1 <- df_schoof1
partijenstemgedrag$schoof1$originvotesrutteschoof1 <- df_schoof1_ruw

```

### Stats!

First, we download the stats files

```{r}

stats_balkenende4_df <- as.data.frame(readxl::read_excel("C:/Users/niels/Downloads/RMSCS/Year 2/Social Networks/labjournal_social_networks/data/partijgedrag_data.xlsx", sheet = 2))

stats_rutte1_df <- as.data.frame(readxl::read_excel("C:/Users/niels/Downloads/RMSCS/Year 2/Social Networks/labjournal_social_networks/data/partijgedrag_data.xlsx", sheet = 4))

stats_rutte2_df <- as.data.frame(readxl::read_excel("C:/Users/niels/Downloads/RMSCS/Year 2/Social Networks/labjournal_social_networks/data/partijgedrag_data.xlsx", sheet = 6))

stats_rutte3_df <- as.data.frame(readxl::read_excel("C:/Users/niels/Downloads/RMSCS/Year 2/Social Networks/labjournal_social_networks/data/partijgedrag_data.xlsx", sheet = 8))

stats_rutte4_df <- as.data.frame(readxl::read_excel("C:/Users/niels/Downloads/RMSCS/Year 2/Social Networks/labjournal_social_networks/data/partijgedrag_data.xlsx", sheet = 10))

stats_schoof1_df <- as.data.frame(readxl::read_excel("C:/Users/niels/Downloads/RMSCS/Year 2/Social Networks/labjournal_social_networks/data/partijgedrag_data.xlsx", sheet = 12))

```

Subsequently, we clean up the stats dataframes and make them ready

```{r, warning=F}

rownames(stats_balkenende4_df) <- stats_balkenende4_df[,1]

stats_balkenende4_df <- stats_balkenende4_df[,-1]

stats_balkenende4_df$coalitie <- as.factor(stats_balkenende4_df$coalitie)
stats_balkenende4_df$oppositie <- as.factor(stats_balkenende4_df$oppositie)

stats_balkenende4_df$ch_ideologie <- as.numeric(stats_balkenende4_df$ch_ideologie)


```

```{r, warning=F}

rownames(stats_rutte1_df) <- stats_rutte1_df[,1]

stats_rutte1_df <- stats_rutte1_df[,-1]

stats_rutte1_df$coalitie <- as.factor(stats_rutte1_df$coalitie)
stats_rutte1_df$oppositie <- as.factor(stats_rutte1_df$oppositie)

stats_rutte1_df$ch_ideologie <- as.numeric(stats_rutte1_df$ch_ideologie)


```

```{r, warning=F}

rownames(stats_rutte2_df) <- stats_rutte2_df[,1]

stats_rutte2_df <- stats_rutte2_df[,-1]

stats_rutte2_df$coalitie <- as.factor(stats_rutte2_df$coalitie)
stats_rutte2_df$oppositie <- as.factor(stats_rutte2_df$oppositie)

stats_rutte2_df$ch_ideologie <- as.numeric(stats_rutte2_df$ch_ideologie)


```

```{r, warning=F}

rownames(stats_rutte3_df) <- stats_rutte3_df[,1]

stats_rutte3_df <- stats_rutte3_df[,-1]

stats_rutte3_df$coalitie <- as.factor(stats_rutte3_df$coalitie)
stats_rutte3_df$oppositie <- as.factor(stats_rutte3_df$oppositie)

stats_rutte3_df$ch_ideologie <- as.numeric(stats_rutte3_df$ch_ideologie)

```

```{r, warning=F}

rownames(stats_rutte4_df) <- stats_rutte4_df[,1]

stats_rutte4_df <- stats_rutte4_df[,-1]

stats_rutte4_df$coalitie <- as.factor(stats_rutte4_df$coalitie)
stats_rutte4_df$oppositie <- as.factor(stats_rutte4_df$oppositie)

stats_rutte4_df$ch_ideologie <- as.numeric(stats_rutte4_df$ch_ideologie)

```

```{r, warning=F}

rownames(stats_schoof1_df) <- stats_schoof1_df[,1]

stats_schoof1_df <- stats_schoof1_df[,-1]

stats_schoof1_df$coalitie <- as.factor(stats_schoof1_df$coalitie)
stats_schoof1_df$oppositie <- as.factor(stats_schoof1_df$oppositie)

stats_schoof1_df$ch_ideologie <- as.numeric(stats_schoof1_df$ch_ideologie)

```

Fill in the slots for votes!

```{r}

#balkenende4
partijenstemgedrag$balkenende4$statsbalk <- stats_balkenende4_df

#Rutte1
partijenstemgedrag$rutte1$statsrutte1 <- stats_rutte1_df

#Rutte2
partijenstemgedrag$rutte2$statsrutte2 <- stats_rutte2_df

#Rutte3
partijenstemgedrag$rutte3$statsrutte3 <- stats_rutte3_df

#Rutte4
partijenstemgedrag$rutte4$statsrutte4 <- stats_rutte4_df

#Schoof1
partijenstemgedrag$schoof1$statsschoof1 <- stats_schoof1_df

```

### Adjacency matrixes

First, we create the adjacancy matrixes

```{r}

#Turns out that R won't make an igraph (properly) if there are NAs in a matrix? Which is why I remove them. 

adj_balkenende4 <- as.matrix(df_balkenende4) %>%
  replace(is.na(.), 0)

adj_rutte1 <- as.matrix(df_rutte1) %>%
  replace(is.na(.), 0)

adj_rutte2 <- as.matrix(df_rutte2) %>%
  replace(is.na(.), 0)

adj_rutte3 <- as.matrix(df_rutte3) %>%
  replace(is.na(.), 0)

adj_rutte4 <- as.matrix(df_rutte4) %>%
  replace(is.na(.), 0)

adj_schoof1 <- as.matrix(df_schoof1) %>%
  replace(is.na(.), 0)

```

Then, we fill them in!

```{r}

#balkenende4
partijenstemgedrag$balkenende4$adjvotesbalk <- adj_balkenende4

#Rutte1
partijenstemgedrag$rutte1$adjvotesrutte1 <- adj_rutte1

#Rutte2
partijenstemgedrag$rutte2$adjvotesrutte2 <- adj_rutte2

#Rutte3
partijenstemgedrag$rutte3$adjvotesrutte3 <- adj_rutte3

#Rutte4
partijenstemgedrag$rutte4$adjvotesrutte4 <- adj_rutte4

#Schoof1
partijenstemgedrag$schoof1$adjvotesschoof1 <- adj_schoof1

```

### Graph

Now you might be thinking, this object was not previously mentioned in the list. Even though you said said it was a good overview of everything that was going to be added on later........Okay I lied, I added this section later on during the process but thought it was more efficient to just add this here :)

First, we create the graphs

```{r}
g_balkenende4 <- graph_from_adjacency_matrix(adj_balkenende4, mode=c("directed"), weighted = F, add.rownames = NA)

g_rutte1 <- graph_from_adjacency_matrix(adj_rutte1, mode=c("directed"), weighted = F, add.rownames = NA)

g_rutte2 <- graph_from_adjacency_matrix(adj_rutte2, mode=c("directed"), weighted = F, add.rownames = NA)

g_rutte3 <- graph_from_adjacency_matrix(adj_rutte3, mode=c("directed"), weighted = F, add.rownames = NA)

g_rutte4 <- graph_from_adjacency_matrix(adj_rutte4, mode=c("directed"), weighted = F, add.rownames = NA)

g_schoof1 <- graph_from_adjacency_matrix(adj_schoof1, mode = "directed", weighted = F, add.rownames = NA)

```

Then, we ~~fill~~ .... *add* them in...

```{r}

partijenstemgedrag$balkenende4$graphbalk <- g_balkenende4
partijenstemgedrag$rutte1$graphrutte1 <- g_rutte1
partijenstemgedrag$rutte2$graphrutte2 <- g_rutte2
partijenstemgedrag$rutte3$graphrutte3 <- g_rutte3
partijenstemgedrag$rutte4$graphrutte4 <- g_rutte4
partijenstemgedrag$schoof1$graphschoof1 <- g_schoof1


```

### Weights

Let's add weights for whether co-voting is smaller than 50%, between 50%-80% or 80%-100%.

```{r}

balkenendeweights1 <- df_balkenende4_ruw %>%
  replace(is.na(.), FALSE) %>%
  lapply(function(x) ifelse(x>=.5, TRUE, FALSE)) %>%
  as.data.frame()

balkenendeweights2 <- df_balkenende4_ruw %>%
  replace(is.na(.), FALSE) %>%
  lapply(function(x) ifelse(x>=.8, TRUE, FALSE)) %>%
  as.data.frame()

```

```{r}

rutte1weights1 <- df_rutte1_ruw %>%
  replace(is.na(.), FALSE) %>%
  lapply(function(x) ifelse(x>=.5, TRUE, FALSE)) %>%
  as.data.frame()

rutte1weights2 <- df_rutte1_ruw %>%
  replace(is.na(.), FALSE) %>%
  lapply(function(x) ifelse(x>=.8, TRUE, FALSE)) %>%
  as.data.frame()

```

```{r}

rutte2weights1 <- df_rutte2_ruw %>%
  replace(is.na(.), FALSE) %>%
  lapply(function(x) ifelse(x>=.5, TRUE, FALSE)) %>%
  as.data.frame()

rutte2weights2 <- df_rutte2_ruw %>%
  replace(is.na(.), FALSE) %>%
  lapply(function(x) ifelse(x>=.8, TRUE, FALSE)) %>%
  as.data.frame()

```

```{r}

rutte3weights1 <- df_rutte3_ruw %>%
  replace(is.na(.), FALSE) %>%
  lapply(function(x) ifelse(x>=.5, TRUE, FALSE)) %>%
  as.data.frame()

rutte3weights2 <- df_rutte3_ruw %>%
  replace(is.na(.), FALSE) %>%
  lapply(function(x) ifelse(x>=.8, TRUE, FALSE)) %>%
  as.data.frame()

```

```{r}

rutte4weights1 <- df_rutte4_ruw %>%
  replace(is.na(.), FALSE) %>%
  lapply(function(x) ifelse(x>=.5, TRUE, FALSE)) %>%
  as.data.frame()

rutte4weights2 <- df_rutte4_ruw %>%
  replace(is.na(.), FALSE) %>%
  lapply(function(x) ifelse(x>=.8, TRUE, FALSE)) %>%
  as.data.frame()

```

```{r}

schoof1weights1 <- df_schoof1_ruw %>%
  replace(is.na(.), FALSE) %>%
  lapply(function(x) ifelse(x>=.5, TRUE, FALSE)) %>%
  as.data.frame()

schoof1weights2 <- df_schoof1_ruw %>%
  replace(is.na(.), FALSE) %>%
  lapply(function(x) ifelse(x>=.8, TRUE, FALSE)) %>%
  as.data.frame()

```

Let's put them in the list as well!

```{r}
#balkenende4
partijenstemgedrag$balkenende4$weights1balk <- balkenendeweights1
partijenstemgedrag$balkenende4$weights2balk <- balkenendeweights2

#Rutte1
partijenstemgedrag$rutte1$weights1rutte1 <- rutte1weights1
partijenstemgedrag$rutte1$weights2rutte1 <- rutte1weights2

#Rutte2
partijenstemgedrag$rutte2$weights1rutte2 <- rutte2weights1
partijenstemgedrag$rutte2$weights2rutte2 <- rutte2weights2

#Rutte3
partijenstemgedrag$rutte3$weights1rutte3 <- rutte3weights1
partijenstemgedrag$rutte3$weights2rutte3 <- rutte3weights2

#Rutte4
partijenstemgedrag$rutte4$weights1rutte4 <- rutte4weights1
partijenstemgedrag$rutte4$weights2rutte4 <- rutte4weights2

#Schoof1
partijenstemgedrag$schoof1$weights1schoof1 <- schoof1weights1
partijenstemgedrag$schoof1$weights2schoof1 <- schoof1weights2

```

### Final save

Saving the list for potential transfer/download

```{r}
saveRDS(partijenstemgedrag, "C:/Users/niels/Downloads/RMSCS/Year 2/Social Networks/labjournal_social_networks/data/partijenstemgedragdatalist.Rdata")

```
