---
title: "Week 2"
author: "Niels Alderliesten"
date: "2026-09-07"
output: html_document
bibliography: references.bib
---

```{r, globalsettings, echo=FALSE, warning=FALSE}
library(knitr)
library(styler)

knitr::opts_chunk$set(echo = TRUE)
opts_chunk$set(tidy.opts=list(width.cutoff=100),tidy=TRUE, warning = FALSE, message = FALSE,comment = "#>", cache=TRUE, class.source=c("test"), results='hold', class.output=c("test2"), cache.lazy= F, eval=F)
options(width = 100)
rgl::setupKnitr()

```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy(position = c('top', 'right'))
klippy::klippy(color = 'lightblue')
klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done')

```

# Week 2

:::: columns
::: {.column width="50%"}
Last compiled on `r format(Sys.time(), '%B, %Y')`

## Descriptives Time!

Installing libraries first ofcourse

```{r}
#install.packages("RSiena")
#install.packages("igraph")

library(RSiena)
library(igraph)
library(tidyverse)
library(igraph)


```

<!-- Viewing random stuff -->

```{r, include=F}
?s501
str(s501)
View(s501)
class(s501)
safetymatrix <- s501

```

```{r, include=F}

#install.packages("network")
#install.packages("sna")
library(network)
library(sna)

library(RSiena)
net1 <- network::as.network(safetymatrix)
drink <- s50a

View(drink)

class(net1)
#snam1 <- sna::nacf(net1, drink[,1])
#Above code was not finished. 

```

## Updated Research Questions

After discussion with Jochem Tolsma in class, I have decided to further venture into the available data of party co-voting.

Based on this data, an interest was established in creating my own network where co-voting percentage on motions could be translated into (stronger or weaker) ties.

The research questions, as well as accompanying papers, were therefore re-newed.

### Research questions

With this new (soon to be made) network structure, the following research questions were formulated:

Descriptive question: *To what extent is there segregation in co-voting among political parties? How has this developed in the past 20 years?*

Explanatory question: *To what extent can (development in) segregation in co-voting be attributed to ideological distance?*

### Papers

The following two papers were selected to aid in answering the above formulated questions.

**"Twitter and divides in the Dutch parliament: social and political segregation in the following, \@-mentions and retweets networks"** [@tolsma2024]

**"Polarisation in Dutch Parliamentary Voting: A Social Network Approach to Polarisation"** (Van der Wal)

## Attribute list

Data that might be useful:

-   Tweede Kamer Open Data

    <https://www.openkamer.org/stats/stemmingen/partijen/>

```{r}
library(rjson)
#openkamerdata <- fromJSON(file = "openkamer-2026-09-13.json")

```

-   The Dutch Parliamentary Behaviour Dataset:

    <https://research.rug.nl/en/publications/the-dutch-parliamentary-behaviour-dataset/>

```{r, results=TRUE}
#library(dutchparl)
#library(R.filesets)


#votes <- loadRDS("votes.rds") 
#print(votesame(examplevotes))

```

Attributes:

-   Political party ID
-   Co-voting score of Party X with Party Y
    -   Longitudinal
    -   Including time point
    -   Differentiation in topics?
-   Ideological score of political parties measured
    -   Longitudinal
    -   Including time point
    -   Differentiation in topics?
-   Coalition/Opposition status
    -   Longitudinal
    -   Including time point

## Chapter 6 SNASS

In chapter 6 of [@SNASS], we first start with analyzing *segregation*.

```{r}
library(RSiena)
library(network)

friend.data.w1 <- s501
friend.data.w2 <- s502
friend.data.w3 <- s503

drink <- s50a
smoke <- s50s

class(friend.data.w1) 
```

As you can see, they are currently a matrix. So I made them a network class!

```{r}

net1 <- network::as.network(friend.data.w1)
net2 <- network::as.network(friend.data.w2)
net3 <- network::as.network(friend.data.w3)
class(net1) #See! They're a network now.

```

I then use the package sna to calculate Moran's I. 'nacf' computes the sample network covariance/correlation for a specified variable. It is also possible to compute Moran's I with this, which I am going to do now.

```{r}
library(sna)

snam1 <- sna::nacf(net1, drink[,1], type = "moran", neighborhood.type = "out", demean=T)
print(snam1) #As you can see, a lot of output given. We need the second list-element. Therefore:
snam1[2] #0.4331431 

```

```{r}

geodistances <- sna::geodist(net1, count.paths = T) #Counts the lengths of a geodesics? among nodes. I think basically it measures the shortest distance
geodistances <- geodistances$gdist
print(geodistances) #As you can see, INF is used for dissconnected nodes. The rest i assume is path length

#First we have to define a weight
weights1 <- geodistances == 1

library(ape)
#ape::Moran.I calculates Moran's I autocorrelation coefficient of variable given a matrix of weights.
ape::Moran.I(drink[,1], scaled = F, weight = weights1, na.rm=T)
#AS you can see, $observed (=0.486134) is close to sna::nacf(0.4331431) but not quite.

```

```{r}
#import function of Jochem":

fMoran.I <- function(x, weight, scaled = FALSE, na.rm = FALSE, alternative = "two.sided", rowstandardize = TRUE) {
    if (rowstandardize) {
        if (dim(weight)[1] != dim(weight)[2])
            stop("'weight' must be a square matrix")
        n <- length(x)
        if (dim(weight)[1] != n)
            stop("'weight' must have as many rows as observations in 'x'")
        ei <- -1/(n - 1)
        nas <- is.na(x)
        if (any(nas)) {
            if (na.rm) {
                x <- x[!nas]
                n <- length(x)
                weight <- weight[!nas, !nas]
            } else {
                warning("'x' has missing values: maybe you wanted to set na.rm = TRUE?")
                return(list(observed = NA, expected = ei, sd = NA, p.value = NA))
            }
        }
        ROWSUM <- rowSums(weight)
        ROWSUM[ROWSUM == 0] <- 1
        weight <- weight/ROWSUM
        s <- sum(weight)
        m <- mean(x)
        y <- x - m
        cv <- sum(weight * y %o% y)
        v <- sum(y^2)
        obs <- (n/s) * (cv/v)
        if (scaled) {
            i.max <- (n/s) * (sd(rowSums(weight) * y)/sqrt(v/(n - 1)))
            obs <- obs/i.max
        }
        S1 <- 0.5 * sum((weight + t(weight))^2)
        S2 <- sum((apply(weight, 1, sum) + apply(weight, 2, sum))^2)
        s.sq <- s^2
        k <- (sum(y^4)/n)/(v/n)^2
        sdi <- sqrt((n * ((n^2 - 3 * n + 3) * S1 - n * S2 + 3 * s.sq) - k * (n * (n - 1) * S1 - 2 * n *
            S2 + 6 * s.sq))/((n - 1) * (n - 2) * (n - 3) * s.sq) - 1/((n - 1)^2))
        alternative <- match.arg(alternative, c("two.sided", "less", "greater"))
        pv <- pnorm(obs, mean = ei, sd = sdi)
        if (alternative == "two.sided")
            pv <- if (obs <= ei)
                2 * pv else 2 * (1 - pv)
        if (alternative == "greater")
            pv <- 1 - pv
        list(observed = obs, expected = ei, sd = sdi, p.value = pv)
    } else {
        if (dim(weight)[1] != dim(weight)[2])
            stop("'weight' must be a square matrix")
        n <- length(x)
        if (dim(weight)[1] != n)
            stop("'weight' must have as many rows as observations in 'x'")
        ei <- -1/(n - 1)
        nas <- is.na(x)
        if (any(nas)) {
            if (na.rm) {
                x <- x[!nas]
                n <- length(x)
                weight <- weight[!nas, !nas]
            } else {
                warning("'x' has missing values: maybe you wanted to set na.rm = TRUE?")
                return(list(observed = NA, expected = ei, sd = NA, p.value = NA))
            }
        }
        # ROWSUM <- rowSums(weight) ROWSUM[ROWSUM == 0] <- 1 weight <- weight/ROWSUM
        s <- sum(weight)
        m <- mean(x)
        y <- x - m
        cv <- sum(weight * y %o% y)
        v <- sum(y^2)
        obs <- (n/s) * (cv/v)
        if (scaled) {
            i.max <- (n/s) * (sd(rowSums(weight) * y)/sqrt(v/(n - 1)))
            obs <- obs/i.max
        }
        S1 <- 0.5 * sum((weight + t(weight))^2)
        S2 <- sum((apply(weight, 1, sum) + apply(weight, 2, sum))^2)
        s.sq <- s^2
        k <- (sum(y^4)/n)/(v/n)^2
        sdi <- sqrt((n * ((n^2 - 3 * n + 3) * S1 - n * S2 + 3 * s.sq) - k * (n * (n - 1) * S1 - 2 * n *
            S2 + 6 * s.sq))/((n - 1) * (n - 2) * (n - 3) * s.sq) - 1/((n - 1)^2))
        alternative <- match.arg(alternative, c("two.sided", "less", "greater"))
        pv <- pnorm(obs, mean = ei, sd = sdi)
        if (alternative == "two.sided")
            pv <- if (obs <= ei)
                2 * pv else 2 * (1 - pv)
        if (alternative == "greater")
            pv <- 1 - pv
        list(observed = obs, expected = ei, sd = sdi, p.value = pv)
    }
}  

fMoran.I(drink[,1], scaled=F, weight=weights1, na.rm = T, rowstandardize = F)
#Now the $observed (0.4331431) is the same! But the primarily change with JOchem's function is that it automatically standardizes. So rowstandardization matters! (nacff does it automatically, ape does not.)

#I don't quite understand the rowstandardization yet.

```

But now, let's calculate a correlation between ctor i and all the alters to whom it is connected (direct or indirectly) and where alters at a larger distances (longer shortest path lengths) are weighted less.

```{r}

#step 1 for each acter i determine the distances (shortest path lengths) to all other nodes in the network.
#Set the distance to yourself as INf.
diag(geodistances) <- Inf

#Step 2: based on these distances decide on how we want to weigh. That is, determine a distance decay function. This one is pretty standard in the spatial autocorrelation literature but actually pretty arbitrary
weights2 <- exp(-geodistances)

#Step 3: decide whether or not we want to row-standardize. You do not:
fMoran.I(drink[,1], scaled=F, weight=weights2, na.rm = T, rowstandardize = F)
# 0.2817188. Apparently, we are a like to alters very close by (path length one) and less so (or even dissimilar) to alters further off.

```

Time to test significance. Let's first generate random graphs. I could not get the code of 6.2 to work sadly. Therefore, I'm trying it out with Zachary

```{r}
#My own question: How to count dyad and triad configurations?

igraph::edge_density(zachary_g, loops=F)
density <- round(edge_density(zachary_g), 2)
trial <- 1000
trialclus <- triallen <- rep(NA, trial)

for (i in 1:trial) {
  random_graph <- sample_gnp(n = 34, p=density, directed=F) #generates the random graphs
  triallen[i] <- mean_distance(random_graph, unconnected=T) #calculates the average path length
  trialclus[i] <- transitivity(random_graph, isolates = c("NaN")) #calculates the clustering
}

par(mfrow = c(1,2))
{
  hist(trialclus, xlim = c(0.0, 0.30), main = "Average clustering coefficient", xlab = "",)
  abline(v = transitivity(zachary_g, isolates=c("NaN")), col = "red", lwd=3)
}

{
  hist(triallen, xlim = c(1.9, 3), main = "Average path length", xlab = "",)
  abline(v = mean_distance(zachary_g, unconnected=T), col = "red", lwd=3)
}


```

As you can see, Zachary has an above expected clustering. But the path length is extremely average.

## Further descriptives

Node descriptives

```{r}
#Example with Zachary from igraph

zachary_g <- make_graph("Zachary")
plot(zachary_g)

#Pretty, now let's make it an adjacancy matrix
zachary_amat <- as_adjacency_matrix(zachary_g, type = "both", sparse = F)
view(zachary_amat)

#NUmber of nodes
igraph::vcount(zachary_g)
#Number of edges
igraph::ecount(zachary_g)
#degree
igraph::degree(zachary_g)
  #pretty plot?
  hist(degree(zachary_g), xlabs='indegree', mains= 'Histogram of indegree')
  #Doesn't work quite yet I think

#Transitivity
igraph::transitivity(zachary_g, type = c("localundirected"), isolates = c("NaN", "zero"))

#Betweenness
igraph::betweenness(zachary_g, directed=F)


```

As you can see. Node 1, 33 and 34 have an extremely high degree, betweenness, but do not have a big transitivity index (which is logical if they are connected with a lot of people, there is bound to be less triads closed)

Diad & Triad descriptives

```{r}
igraph::dyad_census(zachary_g) #Would give the number of relationships, including whether it is mutual, assymmetric or non-existent. Therefore only works with directed graphs, which zachary_g is not :). It therefore only gives mutual and null, not asymmetical etc.

igraph::triad_census(zachary_g) #Same problem as with dyad_census. BUt it still gives the amount of connections.
sna::triad.census(zachary_amat)

#Either
45 / (393+45) #= 0.1027397
45 / (3971 + 1575 + 393) #= 0.007577033
igraph::transitivity(zachary_g, type = c("global"), isolates = c("NaN", "zero")) # 0.2556818
sna::gtrans(zachary_amat) #0.2556818
#Conclusion: I do not know how to calculate transitivity

igraph::transitivity(zachary_g, type = "global")
sna::gtrans(zachary_amat)

triad_g <- data.frame(sna::triad.census(zachary_amat))

transitivity_g <- (3 * triad_g$X300)/(triad_g$X201 + 3 * triad_g$X300)
transitivity_g

```

Network Visualisation

```{r}
plot(zachary_g)
#Let's make those with a larger betweeness, also larger in the graph

V(zachary_g)$size = betweenness(zachary_g, normalized = T, directed = F) * 70 + 15 #*70 refers to the size of the larger betweeness (or something), while the +15 refers to the size of the other nodes
plot(zachary_g)

#Make them further apart!
set.seed(2345)
layout1 <- layout_with_mds(zachary_g)
plot(zachary_g, layout=layout1)

#We change the coordinates of the nodes in the new layout we created
layout1
layout1[1, 1] <- 4 #THe x axis of node 1 is now further to the right
layout1[34, 1] <- -4
plot(zachary_g, layout=layout1)

#Let's add a note for transparancy sake (as we altered the coordination of two nodes)
#Is this really the same as statistically changing something? 
plot(zachary_g, layout=layout1, margin = c(0,0,0,0))
legend(x = -2, y = -1.5, c("Note: the position of nodes 1 and 34 have been altered by Niels Alderliesten \n and
                         Jochem Toslam for visualisation purposes only. These changes do not \n
                         reflect network properties"), bty = "n", cex =0.8) #For some reason the legend does not show up

```
:::




::::
