---
title: "Week4"
author: "Niels Alderliesten"
date: "2026-09-21"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r, globalsettings, warning=FALSE}
library(knitr)
library(styler)

knitr::opts_chunk$set(echo = TRUE)
opts_chunk$set(tidy.opts=list(width.cutoff=100),tidy=TRUE, warning = FALSE, message = FALSE,comment = "#>", cache=TRUE, class.source=c("test"), results='hold', class.output=c("test2"), cache.lazy= F, eval=F)
options(width = 100)
rgl::setupKnitr()

```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy(position = c('top', 'right'))
klippy::klippy(color = 'lightblue')
klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done')

```

## Necessary libraries

```{r}
library(RSiena)
library(igraph)
library(tidyverse)
library(igraph)
```

## Network Visualisation and Webscraping!

Today, we will be learning how to make pretty graphs/visualisations (and
perhaps webscraping)

In order to make a good visualisation, it is necessary to know
beforehand what you want to illustrate. In order to so, you must
ofcourse know your data and your theory. Plot first before running
RSiena. LOOK first to see if they are relevant. Afterwards, when you
have your results and your conclusions, then make a graph to illustrate
this to the readers.

## Network Visualistions

In the following section, one will see network visualisations and
descriptives that were already (partially)

### Loading it in.

```{r}

#Example with Zachary from igraph

zachary_g <- make_graph("Zachary")
plot(zachary_g)

#Pretty, now let's make it an adjacancy matrix

```

``` {r}
zachary_amat <- as_adjacency_matrix(zachary_g, type = "both", sparse = F)
view(zachary_amat)

#Let's check, are they actually undirected?
isSymmetric(zachary_amat) 
#Yes they are. 

```

What do we know based on this: Total 34 nodes. Undirected Ties.

```{r}
#NUmber of nodes
igraph::vcount(zachary_g)
#Number of edges
igraph::ecount(zachary_g)
#degree
igraph::degree(zachary_g)
  #pretty plot?
  hist(degree(zachary_g), xlabs='indegree', mains= 'Histogram of indegree')
  #Doesn't work quite yet I think

#Transitivity
igraph::transitivity(zachary_g, type = c("localundirected"), isolates = c("NaN", "zero"))

#Betweenness
igraph::betweenness(zachary_g, directed=F)

```

As you can see. Node 1, 33 and 34 have an extremely high degree,
betweenness, but do not have a big transitivity index (which is logical
if they are connected with a lot of people, there is bound to be less
triads closed)

### Diad & Triad descriptives

```{r}
igraph::dyad_census(zachary_g) #Would give the number of relationships, including whether it is mutual, assymmetric or non-existent. Therefore only works with directed graphs, which zachary_g is not :). It therefore only gives mutual and null, not asymmetical etc.

igraph::triad_census(zachary_g) #Same problem as with dyad_census. BUt it still gives the amount of connections.
sna::triad.census(zachary_amat)

#Either
45 / (393+45) #= 0.1027397
45 / (3971 + 1575 + 393) #= 0.007577033
igraph::transitivity(zachary_g, type = c("global"), isolates = c("NaN", "zero")) # 0.2556818
sna::gtrans(zachary_amat) #0.2556818
#Conclusion: I do not know how to calculate transitivity

igraph::transitivity(zachary_g, type = "global")
sna::gtrans(zachary_amat)

triad_g <- data.frame(sna::triad.census(zachary_amat))

transitivity_g <- (3 * triad_g$X300)/(triad_g$X201 + 3 * triad_g$X300)
transitivity_g

```

Definition of transitivity is: hoeveel triads die gesloten zijn /
(aantal dat gesloten zijn + gesloten kunnen worden) Why do you have to
do \* 3? Because a triad can be closed for three egos. This is different
than with local transitivity, which measures transitivity on an
ego-level, how connected are the people I am connected with (to
eachother). Ego can have higher local transitivity.

Verder nadenken: Global repricocity werkt hetzelfde (bij directed kan
dat alleen, is wel allemaal moeilijker), waarbij je het \*2 moet doen
i.p.v. 3.

Conclusie dus: 1 en 34 zijn niet met elkaar verbonden. Hebben een hoge
mate van betweenness, maar een lage mate van transitivity. Let's
visualise this!

### Network Visualisation

To alter the visuals/characteristics of graph: V = vertice
(characteristics of nodes) E = edge (characteristics of relations)

```{r}
plot(zachary_g)
#Let's make those with a larger betweeness, also larger in the graph

V(zachary_g)$size = betweenness(zachary_g, normalized = T, directed = F) * 60 + 10 #*70 refers to the size of the larger betweeness (you are weighing the betweenness 60 times more), while the +15 refers to the size of the other nodes, the minimal grootte is 15. 
#Make the size of plots consistent. Define the plot size beforehand. Turn this into a png, then see if it is correct. 
plot(zachary_g)

#Make them further apart!
set.seed(2345)
layout1 <- layout_with_mds(zachary_g)
plot(zachary_g, layout=layout1)

#We change the coordinates of the nodes in the new layout we created
layout1
layout1[1, 1] <- 4 #THe x axis of node 1 is now further to the right
layout1[34, 1] <- -4
plot(zachary_g, layout=layout1)

#Let's add a note for transparancy sake (as we altered the coordination of two nodes)
#Is this really the same as statistically changing something? 
plot(zachary_g, layout=layout1, margin = c(0,0,0,0))
legend(x = -2, y = -1.5, c("Note: the position of nodes 1 and 34 have been altered by Niels Alderliesten \n and
                         Jochem Toslam for visualisation purposes only. These changes do not \n
                         reflect network properties"), bty = "n", cex =0.8) #For some reason the legend does not show up



```

The question now remains, is the visualisation of the network this way
ethical? Or is it data-manipulation? Up to you to decice, gray area.
Images are not neutral, think about that. Either way, disclose it.

## Webscraping

Webscraping, in essence, is gathering/fishing/scraping information from
internet. This can be done automatically or manually. Webscraping is not
the same as Data-sharing also exists, where people can give their
Whatsapp information for instance (who did you message, not content of
message)

**Benefits**:

-   Actual observed behaviour.

-   Cheap for large-scale research.

-   Time/geo stamped.

**Limitations**:

-   More complex analysis (more in-depth data, more difficult it is to
    analyze)

-   Selectivity of posts

-   Sampling

**Ethics**:

-   Is it allowed by the author/owner of data?

-   Is it allowed by the law?

-   Is it allowed by myself? Do I find it ethical?

-   Is it allowed by the ethical commission of your research institute

**APIs**

Aplication Programming Interface: Websites that allow/are designed for
webscraping.

**Crawling** 
What we are going to do.

**Required libaries**
```{r}
######################################### Title: Webscraping in R Author: Bas Hofstra Version:
######################################### 29-07-2021

# start with clean workspace
rm(list = ls())

# install.packages('data.table')
library(data.table)  # mainly for faster data handling
library(tidyverse)  # I assume you already installed this one!
# install.packages('httr') # we don't need this for now require(httr)
install.packages("xml2")
require(xml2)
install.packages("rvest")
require(rvest)
install.packages("devtools")
require(devtools)
# Note we're doing something different here. We're installing a *latest* version directly from
# GitHub This is because the released version of this packages contains some errors!
devtools::install_github("jkeirstead/scholar")

require(scholar)

# define workdirectory, note the double *backslashes* if you're on windows setwd('/yourpathhere)'
```

```{r}
url <- "https://nos.nl/zoeken?q=jochem+tolsma&page=1"

nos <- rvest::read_html(url)

nos

nos %>% rvest::html_elements("main")
#Nu maakt hij een xml nodeset. Hierna gaan we steeds dieper.


lijst <- nos %>% rvest::html_elements("main") %>%
  html_elements("ul") %>% #Want ul was de volgende set.
  html_elements("li") #Dan wil ik nu verder naar de lists, waar die artikelen van de zoekopdracht stonden

lijst[1] %>% html_children() #Onderliggende elements kan je ook m,et children doen. Zoals je kunt zien zit hier weer die <a in. In die <a willen we de href hebben.

lijst[1] %>% html_children %>% html_attr(name="href") #Met attr halen de attributes van dat element eruit, met name kan je weer in die functie specificeren wat je eruit haalt. 


```
Hoe specifieker je wordt, hoe groter de kans dat als de website verandert, je de weg kwijt bent. 


Haal href, datetime attr, en de tekst van datetime eruit. 

```{r}
artikel_partial <- nos %>% rvest::html_elements("main") %>%
  html_elements("ul") %>% 
  html_elements("li") %>%
  html_elements("a")%>%
  html_attr(name="href")

artikellinks <- as.data.frame(paste0("https://nos.nl/", artikel_partial))

artikeldatum <-  nos %>% rvest::html_elements("main") %>%
  html_elements("ul") %>% 
  html_elements("li") %>%
  html_elements("a")%>%
  html_element("time") %>%
  html_attr(name="datetime") %>%
  as.data.frame()

artikeldatumtext <-  nos %>% rvest::html_elements("main") %>%
  html_elements("ul") %>% 
  html_elements("li") %>%
  html_elements("a")%>%
  html_element("time") %>%
  html_text() %>%
  as.data.frame()



```

Laten we het proberen met data van openkamer.org. 

Update: Het adequaat installeren van de .json files was mij helaas niet gelukt.

```{r, include =F}

library(jsonlite)
?jsonlite

openkamer_fulldata <- fromJSON("C:/Users/niels/Downloads/RMSCS/Year 2/Social Networks/labjournal_social_networks/data/openkamer-2026-09-13.json")

openkamer_fulldata_attempt2 <- fromJSON("C:/Users/niels/Downloads/RMSCS/Year 2/Social Networks/labjournal_social_networks/data/openkamer-2026-09-13.json", simplifyVector = T)


?read_json

class(openkamer_fulldata$fields)
View(openkamer_fulldata$fields)

parties <- which(!is.na(openkamer_fulldata$fields$party_name))
parties

View(openkamer_fulldata$fields$party_name)

class(openkamer_fulldata)

```

## Homework: Pretty plot

Behold, my attempt at a pretty plot with my own collected data:

```{r}
plot(g_schoof1)
```
```{r}

edge_density(g_schoof1)
#Density is 0.659

igraph::vcount(g_schoof1)
#Number of nodes = 14

igraph::ecount(g_schoof1)
#Number of edges is 120

#degree
igraph::degree(g_schoof1, mode = c("in"))
#FVD has zero in-degrees
igraph::degree(g_schoof1, mode = c("out"))
#Denk has 11 out-degrees but only 5 indegrees
igraph::degree(g_schoof1, mode = c("total"))
#CU and D66 have high degree. 

#Betweenness
igraph::betweenness(g_schoof1, directed=T)
#Of Note. D66 and CU have incredibly high betweenness score. 

```

As you can see. Node 1, 33 and 34 have an extremely high degree,
betweenness, but do not have a big transitivity index (which is logical
if they are connected with a lot of people, there is bound to be less
triads closed)

### Diad & Triad descriptives

```{r}
igraph::dyad_census(g_schoof1) 

igraph::triad_census(g_schoof1) #Same problem as with dyad_census. BUt it still gives the amount of connections.



igraph::transitivity(g_schoof1, type = "global")
sna::gtrans(adj_schoof1)

triad_g <- data.frame(sna::triad.census(adj_schoof1))
table(triad_g)

E(g_schoof1)$arrow.size = 0.4
V(g_schoof1)$size = 10
plot(g_schoof1)

```

``` {r}

layout2 <- layout_with_mds(g_schoof1)

layout2[1, 1] <- -(0.4)
layout2[2, 1] <- +(2.67)
layout2[3,1] <- + (0.75)
layout2[4,1] <- +(4.84)
layout2[5,1] <- +(4.84)
layout2[6,1] <- +(2.75)
layout2[7,1] <- -(1.11)
layout2[8, 1] <- -(3.5)
layout2[9,1] <- -(0.04)
layout2[10,1] <- +(3)
layout2[11,1] <- -(3.8)
layout2[12,1] <- -(0.5)
layout2[13,1] <- +(1.6)
layout2[14,1] <- -(2.25)

plot(g_schoof1, layout=layout2, vertex.size=15)

V(g_schoof1)$size = degree(g_schoof1, mode="in") *1.5 
plot(g_schoof1, layout=layout2)
V(g_schoof1)$size = degree(g_schoof1, mode="out") *1.5 
plot(g_schoof1, layout=layout2)

clp <- cluster_optimal(g_schoof1)
plot(clp, g_schoof1)

partijenstemgedrag <- loadRDS("C:/Users/niels/Downloads/RMSCS/Year 2/Social Networks/labjournal_social_networks/data/partijenstemgedragdatalist.Rdata")
```

```{r}

clpbalk <- cluster_optimal(g_balkenende4)
plot(clpbalk, g_balkenende4)

clprutte1 <- cluster_optimal(g_rutte1)
plot(clprutte1, g_rutte1)

clprutte2 <- cluster_optimal(g_rutte2)
plot(clprutte2, g_rutte2)

clprutte3 <- cluster_optimal(g_rutte3)
plot(clprutte3, g_rutte3)

clprutte4 <- cluster_optimal(g_rutte4)
plot(clprutte4, g_rutte4)

clp <- cluster_optimal(g_schoof1)
plot(clp, g_schoof1)


```

```{r}

plot(g_balkenende4)
plot(g_rutte1)
plot(g_rutte2)
plot(g_rutte3)
plot(g_rutte4)
plot(g_schoof1)


```

