---
title: "Lab journal Social Networks"
author: "Niels Alderliesten"
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Week 1

<!--- I need to fix the lay-out to hide results etc. --->

::::: columns
::: {.column width="50%"}

Last compiled on `r format(Sys.time(), '%B, %Y')`

This page documents my progress made during my first week following the course Social Networks

<br>

------------------------------------------------------------------------

## Research questions

We were asked to formulate a Descriptive and Explanatory research question. To aid in answering these research questions, we were asked to find 2 research papers relevant to the subject at hand. 
After much deliberating and thinking on what I would like to research (while fitting a social network analysys perspective), I have the following research questions currently in mind:

Descriptive question:
*Do MPs elected in cabinet have a higher in-degree out-degree popularity than non-elected MPs?*

Explanatory question:
*Does inter-party interaction on Twitter influence party collaboration in amendments?*

<!--- this is a reference, stored in references.bib. It will be automatically added at the bottom of the page --->

------------------------------------------------------------------------

##Papers

**"Twitter and divides in the Dutch parliament: social and political segregation in the following, @-mentions and retweets networks"**

**Political Opinions and Offline Social Networks in a Swiss Student Community**

:::



## R Tutorial

We were also asked to read and go through Appendix A and B of *insert reference here*. I am currently still debating on how best to include this in the lab journal. This R tutorial section will therefore be updated on a later date. 

```{r, include=FALSE}
library(tidyverse)
library(haven)

```

##Appendix A

<!--- Random code testing --->
```{r, include=FALSE}
tesvariable <- 4
anothertestvariable <- 6
#look in your environment tab. You now see you have something there. 
ls()
list = ls()
list
ls()
rm(list)
ls()
#? :-)
rm(list=ls())
ls()

?ls

```

<!-- Loading folders, analyzing the data-frame -->

```{r,include=FALSE}

cv08 <- haven::read_sav(file = "Cultural_Changes_2008.sav")
cv10 <- haven::read_sav(file = "Cultural_Changes_2010.sav")

head(names(cv08))
?head
summary(cv08)
head(cv08)
str(cv08, 3)
# what happens if you change the 1 with a 2 or 3? Check the help function or just try and see.
class(cv08)
attributes(cv08)

```

```{r, include=FALSE}
table(cv08$lftop, useNA = "always")
cv08$lftop_new <- cv08$lftop
cv08$lftop_new[cv08$lftop_new == 99] <- NA

table(cv08$lftop_new, useNA = "always")
attributes(cv08$lftop_new)
attr(cv08$lftop, "labels")

```

<!-- Changing educ3 and eslacht to proper labels and values-->


```{r, include=FALSE}
cv08 <- mutate(.data = cv08, lftop_new = na_if(lftop, 99))
table(cv08$lftop_new, useNA = "always")


cv08 %>%
  mutate(
    educ3 = case_match(
      var006n,
      c(-3, -1, 10) ~ "missing",
      c(1) ~ "primary",
      c(2, 3, 4) ~ "secondary",
      c(5, 6, 7, 8) ~ "tertiary"
    ),
    educ3 = as.factor(educ3)
  ) -> cv08

table(cv08$educ3)

class(cv08$geslacht)
attributes(cv08$geslacht)$labels
table(cv08$geslacht, useNA = "always")


cv08 <- cv08 %>%
    mutate(geslacht_n = case_match(geslacht, "9" ~ NA, "M" ~ 0, "V" ~ 1), geslacht_n = labelled(geslacht_n, labels = c(man = 0, vrouw = 1)))

```

<!-- Re-labelling int055 through int057-->

```{r, include=FALSE}
attr(cv08$int055, "labels")
attr(cv08$geslacht_n, "labels")

cv08 <- cv08 %>% 
  mutate(int055_n = case_match(int055, 
                               1 ~ 4, 
                               2 ~ 3, 
                               3 ~ 2, 
                               4 ~ 1, 
                               .default = NA),
         int055 = labelled(int055_n, labels = c('Zeer groot' = 4, 'Groot' = 3, 'Niet zo groot' = 2, 'Helemaal geen tegenstelling' = 1 )))

cv08 <- cv08 %>% 
  mutate(int056_n = case_match(int056, 
                               1 ~ 4, 
                               2 ~ 3, 
                               3 ~ 2, 
                               4 ~ 1, 
                               .default = NA),
         int055 = labelled(int056_n, labels = c('Zeer groot' = 4, 'Groot' = 3, 'Niet zo groot' = 2, 'Helemaal geen tegenstelling' = 1 )))

cv08 <- cv08 %>% 
  mutate(int057_n = case_match(int057, 
                               1 ~ 4, 
                               2 ~ 3, 
                               3 ~ 2, 
                               4 ~ 1, 
                               .default = NA),
         int055 = labelled(int057_n, labels = c('Zeer groot' = 4, 'Groot' = 3, 'Niet zo groot' = 2, 'Helemaal geen tegenstelling' = 1 )))

#Calculating the mean, if amount of NA is smaller than 2

cv08 <- cv08 %>%
  rowwise () %>%
  mutate(int_mean = mean(c(int055_n, int056_n, int057_n), na.rm = T), nmis = sum(c(is.na(int055_n), is.na(int056_n), is.na(int057_n))), int_mean = ifelse(nmis < 2, int_mean, NA)) %>%
  dplyr::select(-nmis)

```

<!-- Selecting the correct variables we want in our new cleaned data-set -->
```{r, include=FALSE}
cv08_cleaned <- cv08 %>%
  dplyr::select(c("we_id", "lftop", "geslacht", "var006n", "v401", "landd"))

cv10_cleaned <- cv10 %>%
  dplyr::select(c("Sleutel", "var002", "var001", "Vltoplop", "V401", "Landd"))

```

```{r, include=FALSE}
names(cv08_cleaned) <- names(cv10_cleaned) <- c("id", "age", "sex", "educ", "health", "region")

cv08_cleaned <- cv08_cleaned %>% mutate(age = na_if(age, 99))
cv10_cleaned <- cv10_cleaned %>% mutate(age = na_if(age, 99))

cv08_cleaned <- cv08_cleaned %>% mutate(
  men = recode_values(sex, "9" ~ -9, "M" ~ 1, "V" ~ 0),
  men = na_if(men, -9), 
  men = labelled(men, c("man" = 1, "vrouw" = 0)))

cv10_cleaned <- cv10_cleaned %>% mutate(
  men = recode_values(sex, c(2) ~ 0), 
  men = labelled(men, c("man" = 1, "vrouw" = 0)))


attr(cv08_cleaned$educ,"labels")



cv08_cleaned <- cv08_cleaned %>%
  mutate(
    educ3 = recode_values(educ, 
                      c(-3,9999999999 ) ~ NA,
                      c(-1, 1)  ~ 1, 
                      c(2,3,4,5) ~ 2, 
                      c(6,7,8) ~ 3),
    educ3 = factor(educ3, 
                   levels = c(1, 2, 3), 
                   labels =  c("primary", "secondary", "tertiary")))

cv10_cleaned <- cv10_cleaned %>%
  mutate(
    educ3 = recode_values(educ, 
                       c(-1, 1) ~ 1,
                       c(2,3,4) ~ 2, 
                       c(5) ~ 3, 
                       c(9999999999) ~ NA),
    educ3 = factor(educ3, 
                   levels = c(1, 2, 3), 
                   labels = c("primary", "secondary", "tertiary")))

cv08_cleaned$wave <- 2008
cv10_cleaned$wave <- 2010

cv08_cleaned$id2 <- rank(cv08_cleaned$id)
cv10_cleaned$id2 <- rank(cv10_cleaned$id)

cv10_cleaned <- cv10_cleaned %>%
  dplyr::select(c("id", "id2", "age", "men", "educ3", "health", "region", "wave"))

#FOllowing code also works btw:
#cv08_cleaned <- cv08_cleaned %>%
#dplyr::select(id, id2, age, men, educ3, health, region, wave)

#Adding them together:
cv08_10_cleaned <- cv08_cleaned %>%
  add_row(cv10_cleaned)

```

<!-- Time to make it into a wide format? -->
``` {r, include=F}
psych::describe(cv08_10_cleaned)

cv08_10_cleaned_wide <- full_join(cv08_cleaned, cv10_cleaned, by = "id2", suffix = c(".2008", ".2010"))

#cv08_10_cleaned_wide <- cv08_10_cleaned_wide %>% 
 # mutate(health = recode_values(health, -1 ~ NA))
#this wasn't working, this is where I stopped :)

```

